---
title: "Introduction to GiottoVisuals"
output: 
  html_document:
    number_sections: true
    toc: true
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Introduction to GiottoVisuals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Description of the module GiottoVisuals

GiottoVisuals contains the main plotting functions of Giotto Suite. 

For full examples of the usage of Giotto visualizations visit http://giottosuite.com

## Script Organization by Prefixes:

aux_ - auxiliary and meta functionality
- aux_output.R - plot output handling
- aux_save.R - plot saving
- aux_defaults.R - set plotting defaults
- aux_visuals.R - general auxiliary plotting functions

plot_ - general plotting functions organized by type of plot
- plot_scatter.R - scatter plots
- plot_violin.R - violin plots
- plot_heatmap.R - heatmaps
- plot_dendrogram.R - dendrograms

vis_ - specific plotting functions organized by topic 
- vis_hvf.R - highly variable features plots
- vis_pc.R - principle components plots
- vis_spatial.R 
- vis_spatial_in_situ.R

gg_ - ggplot2 related
- gg_info_layers.R - modular internals for plotting the different layers of information


Other:
- color_palettes.R - color palettes and factory functions


Package:
- package_imports.R - all imports
- globals.R - global variables
- dd.R - dummy documentation for inheriting
- zzz.R - onloads


## Spatial plots

Load the package

```{r, eval=FALSE}
library(GiottoVisuals)
```

Create a small Giotto object

```{r, eval=FALSE}
expression_matrix <- matrix(rnorm(1000), nrow = 10)
colnames(expression_matrix) <- paste0("cell_", 1:100)
rownames(expression_matrix) <- paste0("gene_", 1:10)

spatial_locations <- data.frame(
    cell_ID = paste0("cell_", 1:100),
    sdimx = rnorm(100),
    sdimy = rnorm(100)
)

giotto_object <- GiottoClass::createGiottoObject(
    expression = expression_matrix,
    spatial_locs = spatial_locations
)
```

Plot the spatial locations

```{r, eval=FALSE}
spatPlot2D(giotto_object)
```


## Session Info

```{r, eval=FALSE}
sessionInfo()
```


```{r, eval=FALSE}
R version 4.5.3 (2026-03-11)
Platform: x86_64-apple-darwin20
Running under: macOS Tahoe 26.4.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-x86_64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] GiottoVisuals_0.2.15

loaded via a namespace (and not attached):
 [1] SummarizedExperiment_1.40.0 gtable_0.3.6                rjson_0.2.23               
 [4] xfun_0.57                   ggplot2_4.0.2               htmlwidgets_1.6.4          
 [7] ggrepel_0.9.8               Biobase_2.70.0              lattice_0.22-9             
[10] vctrs_0.7.3                 tools_4.5.3                 generics_0.1.4             
[13] parallel_4.5.3              stats4_4.5.3                tibble_3.3.1               
[16] colorRamp2_0.1.0            pkgconfig_2.0.3             Matrix_1.7-4               
[19] data.table_1.18.2.1         checkmate_2.3.4             RColorBrewer_1.1-3         
[22] S7_0.2.1                    S4Vectors_0.48.0            lifecycle_1.0.5            
[25] compiler_4.5.3              farver_2.1.2                GiottoUtils_0.2.5          
[28] terra_1.9-11                Seqinfo_1.0.0               codetools_0.2-20           
[31] scattermore_1.2             htmltools_0.5.9             GiottoClass_0.5.1          
[34] yaml_2.3.12                 lazyeval_0.2.2              plotly_4.12.0              
[37] tidyr_1.3.2                 pillar_1.11.1               SingleCellExperiment_1.32.0
[40] DelayedArray_0.36.1         magick_2.9.1                abind_1.4-8                
[43] gtools_3.9.5                tidyselect_1.2.1            digest_0.6.39              
[46] purrr_1.2.2                 dplyr_1.2.1                 labeling_0.4.3             
[49] cowplot_1.2.0               fastmap_1.2.0               grid_4.5.3                 
[52] colorspace_2.1-2            cli_3.6.6                   SparseArray_1.10.10        
[55] magrittr_2.0.5              S4Arrays_1.10.1             withr_3.0.2                
[58] scales_1.4.0                backports_1.5.1             rmarkdown_2.31             
[61] XVector_0.50.0              httr_1.4.8                  matrixStats_1.5.0          
[64] igraph_2.2.2                otel_0.2.0                  reticulate_1.45.0          
[67] png_0.1-9                   SpatialExperiment_1.20.0    evaluate_1.0.5             
[70] knitr_1.51                  GenomicRanges_1.62.1        IRanges_2.44.0             
[73] viridisLite_0.4.3           rlang_1.2.0                 Rcpp_1.1.1                 
[76] glue_1.8.1                  BiocGenerics_0.56.0         rstudioapi_0.18.0          
[79] jsonlite_2.0.0              R6_2.6.1                    MatrixGenerics_1.22.0   
```

